### 2023-06-13 resilience-safe-down.tex
up

Let us choose $\Bad$ as the set of states from which the counter $c_{mutex}$ is not bounded by $k$.
%			at least one counter is unbounded. 
This set is upward-closed by (upward) compatibility in VASS.

As an example, let us choose $\Safe=S \setminus \Bad$. Then, since $\Bad=\uparrow \Bad$, we have: $\Safe=\downarrow \Safe$.

%
% \alain{definir  ideal-effective WSTS}
%
\begin{theorem}\label{up-down}
The resilience problem is decidable for ideal-effective WSTS with 
$\Safe=\downarrow \Safe$
and $\Bad=\uparrow \Bad$
and
the additional hypothesis that
for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
% \alain{est-ce que  $\pred(D)$ is downward closed avec D downward closed serait suffisant ? si l'ordre est omega2 alors closbas+inclusion est wqo donc il faudrait juste que la suite croissante $pred^n(D)+le reste$ converge...}
\end{theorem}

@@

\alain{indecidabilite en enlevant une ou des hypotheses: The resilience problem is undecidable for ideal-effective WSTS}

>>>>

Let us now consider the case $\Bad=\uparrow \Bad$ and $\Safe=\downarrow \Safe$.
It is of interest to note this case can be linked to the problem of mutual exclusion: indeed, one can choose $\Bad$ as the set of states from which the counter $c_{mutex}$ is not bounded by $k$, and $\Safe$ to be the downward-closed complement of $\Bad$. In this case, our contribution consist in the following:

%
\begin{theorem}\label{up-down}
The resilience problem is decidable for ideal-effective WSTS with 
$\Safe=\downarrow \Safe$
and $\Bad=\uparrow \Bad$
and
the additional hypothesis that
for all downward-closed set $D \subseteq S$, the set $\pred^*(D)$ is downward-closed.
% \alain{est-ce que  $\pred(D)$ is downward closed avec D downward closed serait suffisant ? si l'ordre est omega2 alors closbas+inclusion est wqo donc il faudrait juste que la suite croissante $pred^n(D)+le reste$ converge...}
\end{theorem}

@@

% \alain{indecidabilite en enlevant une ou des hypotheses: The resilience problem is undecidable for ideal-effective WSTS}

% \mathieu{
% Home-state indécidable pour les minsky machine (preuve très rapide: M à $2$ compteurs s'arrête ssi M' visite $(0,0,0)$ depuis $\uparrow(0,0,1)$, où M' est M mais avec un troisième compteur qui commence à $1$ et qui simule M jusqu'à ce que M s'arrête et à ce moment là décroit les deux premiers compteurs puis le troisième et atteint $(0,0,0)$).
% or les minsky machine peuvent être simulées par les reset petri nets, on réduit le home-state des minsky machine au home-state des reset petri nets, ce qui donne l'indécidabilité du home-state pour les WSTS en général.}

Without the downward-compatible hypothesis, the problem becomes undecidable. Indeed, it is 
undecidable for Reset 
%Petri-nets
VAS whether zero (the vector containing only zeroes, also denoted~$ \textbf{0}$) is a home-state, where it is a particular case of resilience with $\Safe = \downarrow \textbf{0}$, $\Bad$ complement of $\Safe$. This stems from the fact that it is undecidable for Minsky machines with more than one counter, whether zero is a home-state.

### 2023-06-13 resilience-safe-up.tex
up

We start with the case $\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$. 
It can be viewed as a generalizeation/extension of coverability, namely,
it asks whether for every element of $\Bad$ it is possible to cover an element of the basis of $\Safe$.

@@

As an
example: in a VASS, we may choose $\Safe$ to be the set of states that are not deadlocks, i.e. from which it is always possible to fire a transition. This set of states is upward-closed and $\Safe=S \setminus \Bad=\uparrow \Safe$.\alain{mal rédigé}

@@

Let us now define $K'_{\mathscr{S}}(\Safe,\Bad)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe,\Bad)$.

%	If $k <  \min(K,n)$, then we check every path of length smaller
%	than $\min(K,n)$ from the ideals of the decomposition of $\Bad$ in the completion \alain{à expliquer mieux}.
\end{proof}

@@

Without the completion-post-effective hypothesis or the pred-basis hypothesis, coverability 
becomes undecidable. For instance the coverability problem for strongly increasing well-structured nets is undecidable~\cite{FMP-ic04}.

Coverability reduce
to resilience when $\Bad = \downarrow \Bad$ and  $\Safe = \uparrow \Safe$.
Indeed $s \to^* \uparrow t$ 
reduces to $(\Safe,\Bad)$-resilience for $\Bad = \downarrow s$ and $\Safe = \uparrow t$.
Hence, without the completion-post-effective hypothesis, 
{\sc Resilience} is undecidable.

>>>>

We start with the case $\Bad=\downarrow \Bad$ and $\Safe=\uparrow \Safe$. 
It can be viewed as a generalizeation
 of coverability, namely,
it asks whether for every element of $\Bad$ it is possible to cover an element of the basis of $\Safe$.

@@

\iffalse
As a
example: in a VASS, we may choose $\Safe$ to be the set of states that are not deadlocks, i.e. from which it is always possible to fire a transition. This set of states is upward-closed and $\Safe=S \setminus \Bad=\uparrow \Safe$.\alain{mal rédigé}
\fi

@@

Let us now define $K'_{\mathscr{S}}(\Safe,\Bad)=\max(k'_{i,j} \mid i=1,..,n$ and $j= 1,..m$) and we deduce that  $\mathscr{S}$ is $k$-resilient iff $k \geq K'_{\mathscr{S}}(\Safe,\Bad)$.
%	If $k <  \min(K,n)$, then we check every path of length smaller
%	than $\min(K,n)$ from the ideals of the decomposition of $\Bad$ in the completion \alain{à expliquer mieux}.
\end{proof}

@@

Without the completion-post-effective hypothesis or the pred-basis hypothesis, coverability 
becomes undecidable. For instance the coverability problem for strongly increasing well-structured nets is undecidable~\cite{FMP-ic04}. 
Coverability reduce
to resilience when $\Bad = \downarrow \Bad$ and  $\Safe = \uparrow \Safe$,
as $s \to^* \uparrow t$ 
reduces to $(\Safe,\Bad)$-resilience for $\Bad = \downarrow s$ and $\Safe = \uparrow t$.
Hence, without the completion-post-effective hypothesis, 
{\sc Resilience} itself is undecidable.

### 2023-06-13 resilience.tex
up

@@

Surprinsingly, we have not found decidability results about the resilience problems for WSTS.

%		
We deduce 
% \textcolor{red}{that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed sets.}
% \mathieu{changer ça}
decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed as seen in Section~\ref{synthesis}.

In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
%
%	 and ideals are recursive 
Our result will rely on coverability instead of reachability. We moreover show undecidability when coverability is not decidable in the case $\Safe = \uparrow \Safe$.

@@

{\bf Synthesis of the main results}\label{synthesis}

>>>>

In this section we introduce the concept of resilience, bounded-resilience and $k$-resilience for well structured transition systems. We show that the resilience problems are decidable
for  completion-post-effective $\omega^2$-WSTS with strong compatibility
in the case $\Bad$ is downward-closed and $\Safe $ is upward-closed,
and that resilience is decidable for downward-compatible ideal-effective WSTS
when $\Safe$ is downward-closed and $\Bad $ is upward-closed. 
We show that the three resilience problems are however undecidable for WSTS with
strong compatibility in general.

@@

Surprinsingly, we have not found decidability results about the resilience problems for WSTS.
%		
We deduce 
% \textcolor{red}{that the three resilience problems are decidable for all four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed sets.}
% \mathieu{changer ça}
decidability results for the four pairs of sets $\Safe$ and $\Bad$ downward-closed and upward-closed as seen in Section~\ref{synthesis}.
In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
%
%	 and ideals are recursive 
Our result will rely on coverability instead of reachability. We moreover show undecidability when coverability is not decidable in the case $\Safe = \uparrow \Safe$.

@@

{\bf Synthesis of the main decidability results}\label{synthesis}
